Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web

### 2024-11-25 autovi_paper.qmd
update wording

Determining the best way to implement the interface was difficult. In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework [@webr], which would have allowed the entire application to run directly in the user's browser. However, this approach was not feasible at the time of writing this paper. The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code. A working Emscripten [@zakai2011emscripten] version of this package, which would be required for `webr`, was not available.

@@

The server-side configuration of `autovi.web` is carefully designed to support its functionality. Most required Python libraries, including `pillow` and `NumPy`, are pre-installed on the server. These libraries are integrated into the Shiny application using the `reticulate` package, which provides an interface between R and Python.

@@

In contrast to `autovi`, `autovi.web` does not use the native Python version of `TensorFlow`. Instead, it leverages `TensorFlow.js`, a JavaScript library that allows the execution of machine learning models directly in the browser. This choice enables native browser execution, enhancing compatibility across different user environments, and shifts the computational load from the server to the client-side. `TensorFlow.js` also offers better scalability and performance, especially when dealing with resource-intensive computer vision models on shinyapps.io.

@@

To allow communication between `TensorFlow.js` and other components of the Shiny application, the `shinyjs` R package is used. This package allows calling custom JavaScript code within the Shiny framework. The specialized JavaScript code for initializing `TensorFlow.js` and calling `TensorFlow.js` for VSS prediction is deployed alongside the Shiny application as additional resources.

@@

The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel. There are two example datasets provided by the web application. The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus]. The lineup example uses residuals from a simulated regression model that has a non-linearity issue. We will walk through the lineup example to further demonstrate the workflow of the web application.

@@

The user can select to upload data as either a single set of residuals or a pre-computed lineup of residuals and nulls. Here we illustrate use with lineup example data sets (@fig-autovi-web-setup).  To use the lineup example data, click the "Use Lineup Example" button. The data status will then update to show the number of rows and columns in the dataset, and the CSV type will automatically be selected to the correct option. Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are set automatically (middle image). If the user is working with a custom dataset, these options must be set accordingly. Regardless of the dataset, the user must manually select the label for the true residual plot, whenever a lineup file is provided, as the web application allows assessments without this label. The last step is to click the play button (right image) to start the assessment.

```{r fig-autovi-web-setup, out.width = "100%", fig.cap = 'To begin the workflow for `autovi` using the lineup example dataset, the user clicks the "Use Lineup Example" button (left) to load the example dataset, during which the data status and CSV type will be automatically updated. The target columns are selected automatically, though the user must manually select the label for the true residual plot (middle), as the web application permits assessment without this label. The user initiates the assessment of the lineup example data by clicking the run button (right).'}
magick::image_read_pdf("figures/autovi_web_setup.pdf")
```

>>>>

Determining the best way to implement the backend was difficult. In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework [@webr], which would have allowed the entire application to run directly in the user's browser. However, this approach was not feasible at the time of writing this paper. The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code. A working Emscripten [@zakai2011emscripten] version of this package, which would be required for `webr`, was not available.

@@

The server-side configuration of `autovi.web` is carefully designed to support its functionality. Most required Python libraries, including `pillow` and `numpy`, are pre-installed on the server. These libraries are integrated into the Shiny application using the `reticulate` package, which provides an interface between R and Python.

@@

In contrast to `autovi`, `autovi.web` does not use the native Python version of `TensorFlow`. Instead, it leverages `TensorFlow.js`, a JavaScript library that allows the execution of machine learning models directly in the browser. This choice enables native browser execution, enhancing compatibility across different user environments, and shifts the computational load from the server to the client-side. `TensorFlow.js` also offers better scalability and performance, especially when dealing with resource-intensive computer vision models on the web.

@@

To allow communication between `TensorFlow.js` and other components of the Shiny application, the `shinyjs` R package [@shinyjs] is used. This package allows calling custom JavaScript code within the Shiny framework. The specialized JavaScript code for initializing `TensorFlow.js` and calling `TensorFlow.js` for VSS prediction is deployed alongside the Shiny application as additional resources.

@@

The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel. There are two example datasets provided by the web application. The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus]. The lineup example uses residuals from a simulated regression model that has a non-linearity issue. We walk through the lineup example to further demonstrate the workflow of the web application.

@@

The user can select to upload data as either a single set of residuals and fitted values in a two (or more) column CSV file or a pre-computed lineup of residuals and null datasets in a three (or more) column CSV file (i.e. multiple sets of residuals and fitted values with a column indicating the set label). Here we illustrate use with lineup example data sets (@fig-autovi-web-setup).  To use the lineup example data, click the "Use Lineup Example" button. The data status will then update to show the number of rows and columns in the dataset, and the CSV type will automatically be selected to the correct option. Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are automatically mapped such that the column named as `.fitted` is mapped to fitted values, `.resid` is mapped to residuals and if applicable, `.sample` to labels of the residual set (middle image). If the user is working with a custom dataset, these options must be set accordingly. Whenever a data containing a lineup, the user must manually select the label for the true residual plot, otherwise the web application does not provide all the results. The last step is to click the play button (right image) to start the assessment.

```{r fig-autovi-web-setup, out.width = "100%", fig.cap = 'To begin the workflow for `autovi` using the lineup example dataset, the user clicks the "Use Lineup Example" button (left) to load the example dataset, during which the data status and CSV type will be automatically updated. The user must manually select the label for the true residual plot (middle) to compute further results. The user initiates the assessment of the lineup example data by clicking the run button (right).'}
magick::image_read_pdf("figures/autovi_web_setup.pdf")
```

### 2024-11-25 autovi_paper.tex
update wording

% The following \newcommands are intended for use in the *current*
% document ("protoType"). It is unlikely that authors of papers submitted
% to the Journal will need them.  (Of course if it turns out that you
% do indeed need them, then by all means feel free to make use of them.)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{lipsum}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand\MiKTeX{{\rmfamily M\kern-.05em \textsc{i\kern-.025em K}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand\PracTeX{{\rmfamily P\kern-.05em \textsc{r\kern-.025em a\kern-.025em
c}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{pmboxdraw}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web},
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Automated Residual Plot Assessment with the R Package autovi and
Shiny App autovi.web}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{ANZJS Quarto Template}

\author{
Weihao Li\addressnum{1},
Dianne Cook\addressnum{1},
Emi Tanaka\addressnum{2},
Susan VanderPlas\addressnum{3} and
Klaus Ackermann
\addressnum{1}
}
\affiliation{
Monash University,
The Australian National University and
University of Nebraska
}

@@

\begin{abstract}
Regression software is widely available, and there are many tools for
generating diagnostics and computing conventional residual tests.
However, the advice remains that the analyst should look at the residual
plot to check the fit. Perhaps, one of the reasons is that conventional
tests are too sensitive, strictly resulting in adequate models being
abandoned. Visually assess the strength of patterns in residual plots
requires human effort and suffers from the potential for inconsistent
decisions from different analysts. Using a lineup protocol, where the
residual plot is embedded among null plots, can help to alleviate
inconsistency, but requires even more human effort. This is the type of
task that in today's world might employ a robot to do the tedious work
for a human. Here we describe a new R package that includes a computer
vision model for automated assessment of residual plots, and an
accompanying Shiny app for ease of use. For a user-provided sample of
residuals, it predicts a measure of visual signal strength (VSS) and
provides a suite of supporting information to assist the analyst decide
on the appropriateness their model fit.
\end{abstract}

\keywords{initial data analysis; statistical graphics; data
visualization; visual inference; computer vision; machine
learning; hypothesis testing; regression analysis}

\ack{The author acknowledges that this template is based on the latex
template for Australian and New Zealand Journal of Statistics and
intersperses some text from the original prototype document.
\newline\vspace{1ex}\newline Opinions and attitudes expressed in this
document, which are not explicitly designated as Journal policy, are
those of the author and are \emph{not} necessarily endorsed by the
Journal, its editorial board, its publisher Wiley or by the Australian
Statistical Publishing Association Inc.}

@@

Regression analysis is a widely used statistical modeling technique
widely for data in many fields. There are a vast array of software for
conducting regression modeling and generating diagnostics. The package
\texttt{lmtest} \citep{lmtest} provides a suite of conventional tests.
The \texttt{stats} package \citep{stats} offers standard diagnostic
plots such as residuals vs.~fitted values, quantile-quantile (Q-Q)
plots, and residuals vs.~leverage plots. Packages like \texttt{jtools}
\citep{jtools}, \texttt{olsrr} \citep{olsrr}, \texttt{rockchalk}
\citep{rockchalk}, and \texttt{ggResidpanel} \citep{ggresidpanel}
provide similar graphical diagnostics, often with alternative aesthetics
or interactive features. All of these tools deliver the types of
diagnostic plots outlined in the classical text by
\citet{cook1982residuals}. The \texttt{ecostats} package
\citep{warton_global_2023} incorporates simulation envelopes into
residual plots, while \texttt{DHARMa} \citep{dharma} compares empirical
quantiles (0.25, 0.5, and 0.75) of scaled residuals to their theoretical
counterparts. \texttt{DHARMa} is particularly focused on detecting model
violations such as heteroscedasticity, incorrect functional forms, and
issues specific to generalized linear and mixed-effect models, like
over/under-dispersion. It also includes conventional test annotations to
help avoid misinterpretation.

@@

However, as discussed in \citet{li2024automated}, the lineup protocol
has significant limitations in large-scale applications due dependence
on human labor. Thus, a computer vision model was developed with an
associated statistical testing procedure to automate the assessment of
residual plots. This model takes a residual plot and a vector of
auxiliary variables (such as the number of observations) as inputs and
outputs the predicted visual signal strength (VSS). This strength
estimates the distance between the residual distribution of the fitted
regression model and the reference distribution assumed under correct
model specification.

@@

\subsection{Why use it}\label{sec-why}

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particular for students new to statistical modeling. The
autovi package makes it easier. Here are three examples of residual
plots, which may appear to have structure. According to autovi, the
visual signal strengths (VSS) of these three examples are approximately
(a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being significant
violations of good residuals, but (a) is consistent with a good residual
plot.}

@@

We recommend using the Shiny app \texttt{autovi.web} if encountering
installation problems.

@@

\caption{\label{fig-autovi-diag}Diagram illustrating the infrastructure
of the R package autovi. The modules in green are primary inputs
provided by users. Modules in blue are overridable methods that can be
modified to accommodate users' specific needs. The module in yellow is a
pre-defined non-overridable method. The modules in red are primary
outputs of the package.}

@@

The initial motivation for developing \texttt{autovi} was to create a
convenient interface for sharing the models described and trained in
\citet{li2024automated}. However, recognizing that the classical normal
linear regression model represents a restricted class of models, we
sought to avoid limiting the potential for future extensions, whether by
the original developers or other users. As a result, the package was
designed to function seamlessly with linear regression models with
minimal modification and few required arguments, while also
accommodating other classes of models through partial infrastructure
substitution. This modular and customizable design allows
\texttt{autovi} to handle a wide range of residual diagnostics tasks.

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail on the software's
website \url{https://tengmcing.github.io/autovi/}.

@@

The \texttt{autovi.web} shiny application extends the functionality of
\texttt{autovi} by offering a user-friendly web interface for automated
residual plot assessment. This eliminates the common challenges
associated with software installation, so users can avoid managing
Python environments or handling version requirements for R libraries.
The platform is cross-platform and accessible on various devices and
operating systems, making it suitable even for users without R
programming experience. Additionally, updates are managed centrally,
ensuring that users always have access to the latest features. This
section discusses the implementation based on \texttt{autovi.web}
version 0.1.0.

@@

The interface \texttt{autovi.web} is built using the \texttt{shiny}
\citep{shiny} and \texttt{shinydashboard} \citep{shinydashboard} R
packages. Hosted on the \href{https://www.shinyapps.io}{shinyapps.io}
domain, the application is accessible through any modern web browser.
The R packages \texttt{htmltools} \citep{htmltools} and
\texttt{shinycssloaders} \citep{shinycssloaders} are used to render
markdown documentation in shiny application, and for loading animations
for shiny widgets, respectively.

Determining the best way to implement the interface was difficult. In
our initial planning for \texttt{autovi.web}, we considered implementing
the entire web application using the \texttt{webr} framework
\citep{webr}, which would have allowed the entire application to run
directly in the user's browser. However, this approach was not feasible
at the time of writing this paper. The reason is that one of the R
packages \texttt{autovi} depends on the R package \texttt{splancs}
\citep{splancs}, which uses compiled Fortran code. A working Emscripten
\citep{zakai2011emscripten} version of this package, which would be
required for \texttt{webr}, was not available.

@@

The server-side configuration of \texttt{autovi.web} is carefully
designed to support its functionality. Most required Python libraries,
including \texttt{pillow} and \texttt{NumPy}, are pre-installed on the
server. These libraries are integrated into the Shiny application using
the \texttt{reticulate} package, which provides an interface between R
and Python.

@@

In contrast to \texttt{autovi}, \texttt{autovi.web} does not use the
native Python version of \texttt{TensorFlow}. Instead, it leverages
\texttt{TensorFlow.js}, a JavaScript library that allows the execution
of machine learning models directly in the browser. This choice enables
native browser execution, enhancing compatibility across different user
environments, and shifts the computational load from the server to the
client-side. \texttt{TensorFlow.js} also offers better scalability and
performance, especially when dealing with resource-intensive computer
vision models on shinyapps.io.

@@

To allow communication between \texttt{TensorFlow.js} and other
components of the Shiny application, the \texttt{shinyjs} R package is
used. This package allows calling custom JavaScript code within the
Shiny framework. The specialized JavaScript code for initializing
\texttt{TensorFlow.js} and calling \texttt{TensorFlow.js} for VSS
prediction is deployed alongside the Shiny application as additional
resources.

@@

The workflow of \texttt{autovi.web} is designed to be straightforward,
with numbered steps displayed in each panel. There are two example
datasets provided by the web application. The single residual plot
example uses the \texttt{dino} dataset from the R package
\texttt{datasauRus} \citep{datasaurus}. The lineup example uses
residuals from a simulated regression model that has a non-linearity
issue. We will walk through the lineup example to further demonstrate
the workflow of the web application.

@@

The user can select to upload data as either a single set of residuals
or a pre-computed lineup of residuals and nulls. Here we illustrate use
with lineup example data sets (Figure~\ref{fig-autovi-web-setup}). To
use the lineup example data, click the ``Use Lineup Example'' button.
The data status will then update to show the number of rows and columns
in the dataset, and the CSV type will automatically be selected to the
correct option. Since the example dataset follows the variable naming
conventions assumed by the web application, the columns for fitted
values, residuals, and labels of residual plots are set automatically
(middle image). If the user is working with a custom dataset, these
options must be set accordingly. Regardless of the dataset, the user
must manually select the label for the true residual plot, whenever a
lineup file is provided, as the web application allows assessments
without this label. The last step is to click the play button (right
image) to start the assessment.

@@

\caption{\label{fig-autovi-web-setup}To begin the workflow for
\texttt{autovi} using the lineup example dataset, the user clicks the
``Use Lineup Example'' button (left) to load the example dataset, during
which the data status and CSV type will be automatically updated. The
target columns are selected automatically, though the user must manually
select the label for the true residual plot (middle), as the web
application permits assessment without this label. The user initiates
the assessment of the lineup example data by clicking the run button
(right).}

@@

This paper presents new regression diagnostics software, the R package
\texttt{autovi} and its accompanying web interface, \texttt{autovi.web}.
It addresses a critical gap in the current landscape of statistical
software. While regression tools are widely available, effective and
efficient diagnostic methods have lagged behind, particularly in the
field of residual plot interpretation.

@@

The current version of \texttt{autovi} can be installed from CRAN, and
source code for both packages are available at
\url{https://github.com/TengMCing/autovi} and
\url{https://github.com/TengMCing/autovi_web} respectively. The web
interface is available from \url{autoviweb.netlify.app}.

>>>>

\usepackage{pmboxdraw}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web},
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis, model diagnostics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Automated Residual Plot Assessment with the R Package autovi and
Shiny App autovi.web}

\author{
Weihao Li\addressnum{1},
Dianne Cook\addressnum{1},
Emi Tanaka\addressnum{2},
Susan VanderPlas\addressnum{3} and
Klaus Ackermann\addressnum{1}
}
\affiliation{
Monash University,
The Australian National University and
University of Nebraska
}

@@

\begin{abstract}
Visually assessing residual plots is a common advice for linear model
diagnostics, however this approach requires manual human evaluation and
thereby is not scalable for assessing many models. Human evaluation also
suffer from the potential for inconsistent decisions from different
analysts. Using a lineup protocol, where the residual plot is embedded
among null plots, can help to alleviate inconsistency, but requires even
more human effort. This is the type of task that in today's world might
employ a robot to do the tedious work for a human. Here we describe a
new R package that includes a computer vision model for automated
assessment of residual plots, and an accompanying Shiny app for ease of
use. For a user-provided sample of residuals, it predicts a measure of
visual signal strength (VSS) and provides a suite of supporting
information to assist the analyst decide on the appropriateness their
model fit.
\end{abstract}

\keywords{initial data analysis; statistical graphics; data
visualization; visual inference; computer vision; machine
learning; hypothesis testing; regression analysis; model diagnostics}

@@

Regression analysis is a widely used statistical modeling technique for
data in many fields. There are a vast array of software for conducting
regression modeling and generating diagnostics. The package
\texttt{lmtest} \citep{lmtest} provides a suite of conventional tests.
The \texttt{stats} package \citep{stats} offers standard diagnostic
plots such as residuals vs.~fitted values, quantile-quantile (Q-Q)
plots, and residuals vs.~leverage plots. Packages like \texttt{jtools}
\citep{jtools}, \texttt{olsrr} \citep{olsrr}, \texttt{rockchalk}
\citep{rockchalk}, and \texttt{ggResidpanel} \citep{ggresidpanel}
provide similar graphical diagnostics, often with alternative aesthetics
or interactive features. All of these tools deliver the types of
diagnostic plots outlined in the classical text by
\citet{cook1982residuals}. The \texttt{ecostats} package
\citep{warton_global_2023} incorporates simulation envelopes into
residual plots, while \texttt{DHARMa} \citep{dharma} compares empirical
quantiles (0.25, 0.5, and 0.75) of scaled residuals to their theoretical
counterparts. \texttt{DHARMa} is particularly focused on detecting model
violations such as heteroscedasticity, incorrect functional forms, and
issues specific to generalized linear and mixed-effect models, like
over/under-dispersion. It also includes conventional test annotations to
help avoid misinterpretation.

@@

However, as discussed in \citet{li2024automated}, the lineup protocol
has significant limitations in large-scale applications due to
dependence on human labor. Thus, a computer vision model was developed
with an associated statistical testing procedure to automate the
assessment of residual plots. This model takes a residual plot and a
vector of auxiliary variables (such as the number of observations) as
inputs and outputs the predicted visual signal strength (VSS). This
strength estimates the distance between the residual distribution of the
fitted regression model and the reference distribution assumed under
correct model specification.

@@

\subsection{Motivation for usage}\label{sec-why}

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particularly for students new to statistical modeling.
The \texttt{autovi} package makes it easier. Here are three examples of
residual plots, which may appear to have structure. According to autovi,
the visual signal strengths (VSS) of these three examples are
approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being
significant violations of good residuals, but (a) is consistent with a
good residual plot.}

@@

We recommend using the Shiny app \texttt{autovi.web} if users encounter
installation problems.

@@

\caption{\label{fig-autovi-diag}Diagram illustrating the infrastructure
of the R package \texttt{autovi}. The modules in green are primary
inputs provided by users. Modules in blue are overridable methods that
can be modified to accommodate users' specific needs. The module in
yellow is a pre-defined non-overridable method. The modules in red are
primary outputs of the package.}

@@

The initial motivation for developing \texttt{autovi} was to create a
convenient interface for sharing the models described and trained in
\citet{li2024automated}. However, recognizing that the classical normal
linear regression model represents a restricted class of models, we
sought to avoid limiting the potential for future extensions, whether by
the original developers or other developers. As a result, the package
was designed to function seamlessly with linear regression models with
minimal modification and few required arguments, while also
accommodating other classes of models through partial infrastructure
substitution. This modular and customizable design allows
\texttt{autovi} to handle a wide range of residual diagnostics tasks.

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail on the software's
website.

@@

The \texttt{autovi.web} package extends the functionality of
\texttt{autovi} by offering a user-friendly web interface for automated
residual plot assessment. This eliminates the common challenges
associated with software installation, so users can avoid managing
Python environments or handling version requirements for R libraries.
The platform is cross-platform and accessible on various devices and
operating systems, making it suitable even for users without R
programming experience. Additionally, updates are managed centrally,
ensuring that users always have access to the latest features. This
section discusses the implementation based on \texttt{autovi.web}
version 0.1.0.

@@

The package \texttt{autovi.web} is built using the \texttt{shiny}
\citep{shiny} and \texttt{shinydashboard} \citep{shinydashboard} R
packages. Hosted on the \href{https://www.shinyapps.io}{shinyapps.io}
domain, the application is accessible through any modern web browser.
The R packages \texttt{htmltools} \citep{htmltools} and
\texttt{shinycssloaders} \citep{shinycssloaders} are used to render
markdown documentation in shiny application, and for loading animations
for shiny widgets, respectively.

Determining the best way to implement the backend was difficult. In our
initial planning for \texttt{autovi.web}, we considered implementing the
entire web application using the \texttt{webr} framework \citep{webr},
which would have allowed the entire application to run directly in the
user's browser. However, this approach was not feasible at the time of
writing this paper. The reason is that one of the R packages
\texttt{autovi} depends on the R package \texttt{splancs}
\citep{splancs}, which uses compiled Fortran code. A working Emscripten
\citep{zakai2011emscripten} version of this package, which would be
required for \texttt{webr}, was not available.

@@

The server-side configuration of \texttt{autovi.web} is carefully
designed to support its functionality. Most required Python libraries,
including \texttt{pillow} and \texttt{numpy}, are pre-installed on the
server. These libraries are integrated into the Shiny application using
the \texttt{reticulate} package, which provides an interface between R
and Python.

@@

In contrast to \texttt{autovi}, \texttt{autovi.web} does not use the
native Python version of \texttt{TensorFlow}. Instead, it leverages
\texttt{TensorFlow.js}, a JavaScript library that allows the execution
of machine learning models directly in the browser. This choice enables
native browser execution, enhancing compatibility across different user
environments, and shifts the computational load from the server to the
client-side. \texttt{TensorFlow.js} also offers better scalability and
performance, especially when dealing with resource-intensive computer
vision models on the web.

@@

To allow communication between \texttt{TensorFlow.js} and other
components of the Shiny application, the \texttt{shinyjs} R package
\citep{shinyjs} is used. This package allows calling custom JavaScript
code within the Shiny framework. The specialized JavaScript code for
initializing \texttt{TensorFlow.js} and calling \texttt{TensorFlow.js}
for VSS prediction is deployed alongside the Shiny application as
additional resources.

@@

The workflow of \texttt{autovi.web} is designed to be straightforward,
with numbered steps displayed in each panel. There are two example
datasets provided by the web application. The single residual plot
example uses the \texttt{dino} dataset from the R package
\texttt{datasauRus} \citep{datasaurus}. The lineup example uses
residuals from a simulated regression model that has a non-linearity
issue. We walk through the lineup example to further demonstrate the
workflow of the web application.

@@

The user can select to upload data as either a single set of residuals
and fitted values in a two (or more) column CSV file or a pre-computed
lineup of residuals and null datasets in a three (or more) column CSV
file (i.e.~multiple sets of residuals and fitted values with a column
indicating the set label). Here we illustrate use with lineup example
data sets (Figure~\ref{fig-autovi-web-setup}). To use the lineup example
data, click the ``Use Lineup Example'' button. The data status will then
update to show the number of rows and columns in the dataset, and the
CSV type will automatically be selected to the correct option. Since the
example dataset follows the variable naming conventions assumed by the
web application, the columns for fitted values, residuals, and labels of
residual plots are automatically mapped such that the column named as
\texttt{.fitted} is mapped to fitted values, \texttt{.resid} is mapped
to residuals and if applicable, \texttt{.sample} to labels of the
residual set (middle image). If the user is working with a custom
dataset, these options must be set accordingly. Whenever a data
containing a lineup, the user must manually select the label for the
true residual plot, otherwise the web application does not provide all
the results. The last step is to click the play button (right image) to
start the assessment.

@@

\caption{\label{fig-autovi-web-setup}To begin the workflow for
\texttt{autovi} using the lineup example dataset, the user clicks the
``Use Lineup Example'' button (left) to load the example dataset, during
which the data status and CSV type will be automatically updated. The
user must manually select the label for the true residual plot (middle)
to compute further results. The user initiates the assessment of the
lineup example data by clicking the run button (right).}

@@

This paper presents new regression diagnostics software, the R package
\texttt{autovi} and its accompanying web interface package,
\texttt{autovi.web}. It addresses a critical gap in the current
landscape of statistical software. While regression tools are widely
available, effective and efficient diagnostic methods have lagged
behind, particularly in the field of residual plot interpretation.

@@

The the current version of \texttt{autovi} can be installed from CRAN,
and source code for both packages are available at
\url{https://github.com/TengMCing/autovi}. The web interface is
available from \url{autoviweb.netlify.app}.
